Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-04-24 paper.tex
Replace C_{0.5} with Q_{0.5}

Input images must have appropriately high resolution for edge detection filter
to work correctly. A criterion for ``goodness'' of an image can be expressed
with $C_{\alpha}$ parameter \cite{samarin2023robust}. Let $\alpha \in [0, 1]$
and $f(x)$ be (band-limited) input image ($x \in \mathbb{R}^2$ or
$\mathbb{R}^3$). We define $C_\alpha$ as follows:
\begin{equation}
  \begin{aligned}
    f_0(x) &= f(x) - \langle f(x) \rangle \\
    C_\alpha &= \frac{\int_{-\alpha\omega}^{\alpha\omega} |\hat{f_0}(z)|^2
      dz}{\int_{-\omega}^{\omega} |\hat{f_0}(z)|^2 dz}
  \end{aligned}
\end{equation}
where $\hat{f_0}$ is a Fourier transform of $f_0$ and $\omega$ is a folding
frequency which depends only on resolution of the image. If a criterion
$C_{0.5} > 0.97$ holds then the image is suitable for application of
aforementioned algorithms for calculation of surface functions, otherwise
image's resolution can be improved with the help of interpolation prior to
computations \cite{samarin2023robust}.

@@

\begin{table*}[!htp]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Sample name & Image type & Dimensions (pixels/voxels) & Image resolution ($\mu m$)
    & $C_{0.5}$ \\
    \hline
    Sandstone & XCT &  $500 \times 500 \times 500$ & 2.25 & 0.9703 \\
    Soil & Reconstruction & $500 \times 500 \times 500$ & 0.72 & 0.9431 \\
    Shale & Reconstruction & $160 \times 160 \times 160$ & 2.25 & 0.8912 \\
    Sandstone & SEM &  $1280 \times 869$ & 0.10 & 0.9613 \\
    \hline
  \end{tabular}
  \caption{Summary of samples used for calculation of correlation functions.}
  \label{tab:summary}
\end{table*}
In this section we compute correlation functions for 2- and 3-dimensional porous
media samples. Table \ref{tab:summary} contains information about these
samples. Note that we compute surface correlation functions only for those
samples for which the criterion $C_{0.5} > 0.97$ holds (the two sandstones and
the soil) \cite{samarin2023robust}. Images of the samples are presented on
\cref{fig:samples}. On plots of correlation functions for isotropic samples we
provide a family of curves with the first argument of the correlation function
being fixed. For anisotropic samples we provide heatmaps of correlation
functions where anisotropy can be easily observed.
\begin{figure*}[tp]
  \centering
  \subfigure[Sanstone XCT]{
    \includegraphics[width=0.4\linewidth]{images/15581.png}
    \label{fig:sample-sandstone-xct}}
  \hfill
  \subfigure[Sandstone SEM (different clusters of void phase are shown in color)]{
    \includegraphics[width=0.4\linewidth, frame]{images/sandstone1.png}
    \label{fig:sample-sandstone}}
  \vskip\baselineskip
  \subfigure[Soil]{
    \includegraphics[width=0.4\linewidth]{images/soil.png}
    \label{fig:sample-soil}}
  \hfill
  \subfigure[Shale]{
    \includegraphics[width=0.4\linewidth]{images/shale.png}
    \label{fig:sample-shale}}
  \caption[]{Images of samples used for calculation of higher-order correlation
    functions.}
  \label{fig:samples}
\end{figure*}

@@

Direct comparison against existing modern approaches, namely described by Malmir
et al. \cite{malmir2018} and by Sun et al. \cite{sun2022third}, is not always
straightforward due to algorithmic differences.  In particular, the latter cited
work computes 3-point correlation functions for sphere packings in the
continuous domain; however, on a GPU it takes approximately 28 hours to perform
$2\times10^6$ samplings (see Table 1 in \cite{sun2022third}).  Considering
verification results in \cref{sec:results}, it seems that it would be much
faster to discretize the sphere packing with appropriate $C_{0.5} > 0.97$ and to
utilize our package to assess correlation functions within tens of minutes
without any loss in accuracy.  For $S_3$ and 3-point surface functions, Malmir
et al. reported computation times of a couple of minutes for a $128^3$ voxels 3D
image with $2\times10^5$ samples on CPU (see section IV in \cite{malmir2018}).
In our case, we sample $\approx 31\times10^9$ triangles within 2 sec on CPU, and
less than a second on GPU.  Our implementation of $C_3$ is even more performant
compared to existing counterparts.

>>>>

Input images must have appropriately high resolution for edge detection filter
to work correctly. A criterion for ``goodness'' of an image can be expressed
with $Q_\alpha$ parameter \cite{samarin2023robust}. Let $\alpha \in [0, 1]$
and $f(x)$ be (band-limited) input image ($x \in \mathbb{R}^2$ or
$\mathbb{R}^3$). We define $Q_\alpha$ as follows:
\begin{equation}
  \begin{aligned}
    f_0(x) &= f(x) - \langle f(x) \rangle \\
    Q_\alpha &= \frac{\int_{-\alpha\omega}^{\alpha\omega} |\hat{f_0}(z)|^2
      dz}{\int_{-\omega}^{\omega} |\hat{f_0}(z)|^2 dz}
  \end{aligned}
\end{equation}
where $\hat{f_0}$ is a Fourier transform of $f_0$ and $\omega$ is a folding
frequency which depends only on resolution of the image. If a criterion
$Q_{0.5} > 0.97$ holds then the image is suitable for application of
aforementioned algorithms for calculation of surface functions, otherwise
image's resolution can be improved with the help of interpolation prior to
computations \cite{samarin2023robust}.

@@

\begin{table*}[!htp]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Sample name & Image type & Dimensions (pixels/voxels) & Image resolution ($\mu m$)
    & $Q_{0.5}$ \\
    \hline
    Sandstone & XCT &  $500 \times 500 \times 500$ & 2.25 & 0.9703 \\
    Soil & Reconstruction & $500 \times 500 \times 500$ & 0.72 & 0.9431 \\
    Shale & Reconstruction & $160 \times 160 \times 160$ & 2.25 & 0.8912 \\
    Sandstone & SEM &  $1280 \times 869$ & 0.10 & 0.9613 \\
    \hline
  \end{tabular}
  \caption{Summary of samples used for calculation of correlation functions.}
  \label{tab:summary}
\end{table*}
In this section we compute correlation functions for 2- and 3-dimensional porous
media samples. Table \ref{tab:summary} contains information about these
samples. Note that we compute surface correlation functions only for those
samples for which the criterion $Q_{0.5} > 0.97$ holds (the two sandstones and
the soil) \cite{samarin2023robust}. Images of the samples are presented on
\cref{fig:samples}. On plots of correlation functions for isotropic samples we
provide a family of curves with the first argument of the correlation function
being fixed. For anisotropic samples we provide heatmaps of correlation
functions where anisotropy can be easily observed.
\begin{figure*}[tp]
  \centering
  \subfigure[Sanstone XCT]{
    \includegraphics[width=0.4\linewidth]{images/15581.png}
    \label{fig:sample-sandstone-xct}}
  \hfill
  \subfigure[Sandstone SEM (different clusters of void phase are shown in color)]{
    \includegraphics[width=0.4\linewidth, frame]{images/sandstone1.png}
    \label{fig:sample-sandstone}}
  \vskip\baselineskip
  \subfigure[Soil]{
    \includegraphics[width=0.4\linewidth]{images/soil.png}
    \label{fig:sample-soil}}
  \hfill
  \subfigure[Shale]{
    \includegraphics[width=0.4\linewidth]{images/shale.png}
    \label{fig:sample-shale}}
  \caption[]{Images of samples used for calculation of higher-order correlation
    functions.}
  \label{fig:samples}
\end{figure*}

@@

Direct comparison against existing modern approaches, namely described by Malmir
et al. \cite{malmir2018} and by Sun et al. \cite{sun2022third}, is not always
straightforward due to algorithmic differences.  In particular, the latter cited
work computes 3-point correlation functions for sphere packings in the
continuous domain; however, on a GPU it takes approximately 28 hours to perform
$2\times10^6$ samplings (see Table 1 in \cite{sun2022third}).  Considering
verification results in \cref{sec:results}, it seems that it would be much
faster to discretize the sphere packing with appropriate $Q_{0.5} > 0.97$ and to
utilize our package to assess correlation functions within tens of minutes
without any loss in accuracy.  For $S_3$ and 3-point surface functions, Malmir
et al. reported computation times of a couple of minutes for a $128^3$ voxels 3D
image with $2\times10^5$ samples on CPU (see section IV in \cite{malmir2018}).
In our case, we sample $\approx 31\times10^9$ triangles within 2 sec on CPU, and
less than a second on GPU.  Our implementation of $C_3$ is even more performant
compared to existing counterparts.
